---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# VisionZeroR

<!-- badges: start -->
[![R-CMD-check](https://github.com/kristine-le1/VisionZeroR/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/kristine-le1/VisionZeroR/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of VisionZeroR package is to consolidate and refine vehicle fatality data from five cities recognized by VisionZero as of Fall 2023 - specifically, Boston, Richmond, Denver, Boulder, and NYC. This package transforms the data into an accessible format for users. Additionally, it includes a feature that allows plotting of vehicle fatality rates, adjusted for population size, for any combination of these five cities. This functionality is particularly useful for analyzing and comparing the effectiveness of these cities in reducing vehicle-related fatalities over time under the VisionZero initiative.

## Installation

You can install the development version of VisionZeroR from [GitHub](https://github.com/) with:

```{r eval=FALSE}
#install.packages("devtools")
devtools::install_github("kristine-le1/VisionZeroR", force = TRUE, dependencies = TRUE, build_vignettes = TRUE)
```


## Example

This is a basic example which shows you how to solve a common problem:

```{r setup}
library(VisionZeroR)
library(vdiffr)
```

## VisionZero vehicular fatality data

The vehicular fatality and population data for the VisionZero cities Boston, Richmond, Boulder, Denver, and NYC are aggregated into the `vz_data_2023` data set. 

```{r}
head(vz_data_2023)
```

In what year did each city have the most vehicular fatalities and how many vehicular fatalities were there?

```{r}
library(dplyr)
```

```{r}

max <- vz_data_2023 |> 
    group_by(city) |> 
    summarise(max_fatalities = max(total_fatal))

edit_vz <- vz_data_2023 |> 
    select(1:3)

max_2 <- left_join(max, edit_vz, by = c("max_fatalities" = "total_fatal", "city" = "city"))

head(max_2)

```

## VisionZero vehicular fatality data visualization

Given a character vector of up to 5 elements made up of city names from the list of Boston, Richmond, NYC, Denver, or Boulder, the `vz_graph` function creates a point graph of vehicular fatalities normalized to the population size of the city for the user-given cities.

```{r}

# Graph vehicular fatality data for just Boston
vz_graph("Boston")

# Graph vehicular fatality data for Boston, NYC, and Boulder
vz_graph(c("NYC", "Boston", "Boulder"))

```


You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this.

In that case, don't forget to commit and push the resulting figure files, so they display on GitHub and CRAN.
In that case, don't forget to commit and push the resulting figure files, so they display on GitHub and CRAN.
