---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# VisionZeroR

<!-- badges: start -->
[![R-CMD-check](https://github.com/kristine-le1/VisionZeroR/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/kristine-le1/VisionZeroR/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of VisionZeroR package is to consolidate and refine vehicle fatality data from five cities recognized by VisionZero as of Fall 2023 - specifically, Boston, Richmond, Denver, Boulder, and NYC. This package transforms the data into an accessible format for users. Additionally, it includes a feature that allows plotting of vehicle fatality rates, adjusted for population size, for any combination of these five cities. This functionality is particularly useful for analyzing and comparing the effectiveness of these cities in reducing vehicle-related fatalities over time under the VisionZero initiative.

## Installation

You can install the development version of VisionZeroR from [GitHub](https://github.com/) with:

```{r}
#install.packages("devtools")
devtools::install_github("kristine-le1/VisionZeroR", force = TRUE)

```


## Example

This is a basic example which shows you how to solve a common problem:

```{r setup}
library(VisionZeroR)
```

The `VisionZeroR` package is a package that aggregates and cleans up vehicular fatality data for 5 [VisionZero](https://visionzeronetwork.org/about/what-is-vision-zero/)-recognized cities that have available data as of Fall 2023 (Boston, Richmond, Denver, Boulder, and NYC) into a user-friendly data set. It also provides a function to graph the vehicular fatalities normalized to population size for 1 to 5 of the cities over the years to see how the cities have done with decreasing vehicular fatalities as VisionZero cities.

## VisionZero vehicular fatality data

The vehicular fatality and population data for the VisionZero cities Boston, Richmond, Boulder, Denver, and NYC are aggregated into the `vz_data_2023` data set. 

```{r}
head(vz_data_2023)
```

In what year did each city have the most vehicular fatalities and how many vehicular fatalities were there?

```{r}
library(dplyr)
```

```{r}

library(dplyr)

max <- vz_data_2023 |> 
    group_by(city) |> 
    summarise(max_fatalities = max(total_fatal))

edit_vz <- vz_data_2023 |> 
    select(1:3)

max_2 <- left_join(max, edit_vz, by = c("max_fatalities" = "total_fatal", "city" = "city"))

```



You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this.

In that case, don't forget to commit and push the resulting figure files, so they display on GitHub and CRAN.
In that case, don't forget to commit and push the resulting figure files, so they display on GitHub and CRAN.
