---
title: "Intro to vz_graph"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vz_vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(VisionZeroR)
```

The `VisionZeroR` package is a package that aggregates and cleans up vehicular fatality data for 5 [VisionZero](https://visionzeronetwork.org/about/what-is-vision-zero/)-recognized cities that have available data as of Fall 2023 (Boston, Richmond, Denver, Boulder, and NYC) into a user-friendly data set. It also provides a function to graph the vehicular fatalities normalized to population size for 1 to 5 of the cities over the years to see how the cities have done with decreasing vehicular fatalities as VisionZero cities.

## VisionZero vehicular fatality data

The vehicular fatality and population data for the VisionZero cities Boston, Richmond, Boulder, Denver, and NYC are aggregated into the `vz_data_2023` data set. 

```{r}
head(vz_data_2023)
```

In what year did each city have the most vehicular fatalities and how many vehicular fatalities were there?

```{r}
library(dplyr)
```

```{r}

max <- vz_data_2023 |> group_by(city) |> summarise(max_fatalities = max(total_fatal))

edit_vz <- vz_data_2023 |> select(c(1:3))

max_2 <- left_join(max, edit_vz, by = join_by("max_fatalities" == "total_fatal", "city" == "city"))

head(max_2)
```

## VisionZero vehicular fatality data visualization

Given a character vector of up to 5 elements made up of city names from the list of Boston, Richmond, NYC, Denver, or Boulder, the `vz_graph` function creates a point graph of vehicular fatalities normalized to the population size of the city for the user-given cities.

```{r}

# Graph vehicular fatality data for just Boston
vz_graph("Boston")

# Graph vehicular fatality data for Boston, NYC, and Boulder
vz_graph(c("NYC", "Boston", "Boulder"))

```


